Results MSRC latent class analysis

### 2021-06-07 manuscript/manuscript.Rmd
update results

rename_vars <- c(
CDE7_NEW = "Told someone",
CDE8 = "Will likely attempt",
CDE9 = "Certain fatal",
CDE10 = "Certain lethal",
CDE11 = "Intended to die",
CDE12 = "Could be fixed",
CDE13 = "Attempts",
CDE16 = "Medical attention (DV)",
CDE17 = "NSSI",
AGE = "Age (cov)",
EDUCATION = "Education (cov)",
GENDER = "Gender (cov)",
ANY_SERVICE = "Service (cov)",
METHOD = "Method (DV)",
RACE = "Race (cov)"
)
df_plot$Variable <- ordered(df_plot$Variable, 
                            levels = unique(df_plot$Variable),
                            labels = rename_vars[unique(df_plot$Variable)])

@@

```{r mplusmodel}
aux_cat <- c("Gender", "Race", "Any_Service", "Method")
aux_cont <- c("Age", "Education", "CDE16")
df_anal <- df[c(indicators, aux_cat, aux_cont)]
cat_dict <- lapply(df_anal[aux_cat], levels)
df_anal[aux_cat] <- lapply(df_anal[aux_cat], as.integer)
cat_vars <- c("CDE7_NEW", "CDE8",
              "CDE9", "CDE10", "CDE11", "CDE12", "CDE13", 
              "CDE17")

  # Create dummies for all categories to get by-category comparisons
  df_dum <- df[c(indicators, aux_cat, aux_cont)]
  dum_vars <- aux_cat[sapply(df_dum[aux_cat], function(x){length(unique(na.omit(x)))}) > 2]
  the_opt <- getOption("na.action")
  options(na.action='na.pass')
  dummies <- model.matrix(as.formula(paste0("~ -1 + ", paste0(dum_vars, collapse = " + "))), df_dum,
                          contrasts.arg = lapply(df_dum[dum_vars], contrasts, contrasts=FALSE))

  dummies <- data.frame(dummies, check.names = FALSE)
  options(na.action=the_opt)
  for(i in dum_vars){
    names(dummies) <- gsub(paste0("^", i, "(?!Other)"), "", names(dummies), perl = TRUE)
  }
  dummie_dict <- gsub("^.+?Other", "Other", names(dummies))
  names(dummies) <- gsub("[ /\\(].*", "", names(dummies))
  df_dum <- cbind(df_dum, dummies)
if(run_everything){
  mod_cont <- mplusObject(
    TITLE = paste0("2 class with auxiliary continuous"),
    VARIABLE = paste0("CLASSES = c (2);\n",
                      paste0("CATEGORICAL = \n", paste0(cat_vars, collapse = "\n"), ";\n"),
                      paste0("USEVARIABLES = \n", paste0(indicators, collapse = "\n"), ";\n"),
                      paste0(c("AUXILIARY = \n",
                               paste0(aux_cont, " (BCH)\n"),
                               #paste0(aux_cat, " (DCATEGORICAL)\n"),
                               ";\n"), collapse = "")),
    ANALYSIS = "TYPE IS MIXTURE;\nLRTSTARTS = 0 0 40 10;\nOPTSEED = 637345;\n",
    OUTPUT = "TECH14;",
    SAVEDATA = paste0("FILE IS dat_2_aux_cont.dat;\nSAVE = CPROBABILITIES;"),
    usevariables = names(df_anal),
    rdata = df_anal
  )

  res_cont <- mplusModeler(mod_cont,
                           modelout = paste0("mod_2_aux_cont.inp"),
                           run = 1L)$results

  mod_cat <- mplusObject(
    TITLE = paste0("2 class with auxiliary categorical"),
    VARIABLE = paste0("CLASSES = c (2);\n",
                      paste0("CATEGORICAL = \n", paste0(cat_vars, collapse = "\n"), ";\n"),
                      paste0("USEVARIABLES = \n", paste0(indicators, collapse = "\n"), ";\n"),
                      paste0(c("AUXILIARY = \n",
                               #paste0(aux_cont, " (BCH)\n"),
                               paste0(aux_cat, " (DCATEGORICAL)\n"),
                               ";\n"), collapse = "")),
    ANALYSIS = "TYPE IS MIXTURE;\nLRTSTARTS = 0 0 40 10;\nOPTSEED = 637345;\n",
    OUTPUT = "TECH14;",
    SAVEDATA = paste0("FILE IS dat_2_aux_cat.dat;\nSAVE = CPROBABILITIES;"),
    usevariables = names(df_anal),
    rdata = df_anal
  )

@@

} else {
  res_cont <- readModels("mod_2_aux_cont.out")
  res_cat <- extract_cat_tests(readLines("mod_2_aux_cat.out"))
  res_catdum <- extract_cat_tests(readLines("mod_2_aux_catdum.out"))
  names(res_catdum) <- dummie_dict
}
```

@@

```{r tabcont, include = TRUE}
tab <- res_cont$lcCondMeans$overall
tab$var <- rename_vars[tab$var]
tab$df <- NULL
names(tab)[1] <- "Variable"
tab <- tab[c(grep("(DV)", tab$Variable), grep("(cov)", tab$Variable)), ]
rownames(tab) <- NULL
knitr::kable(tab, digits = 2, caption = "Class differences in continuous auxiliary variables.")
```

```{r tabcat, include = TRUE}
ct_dct <- cat_dict
names(ct_dct) <- toupper(names(ct_dct))

cats <- do.call(rbind, lapply(c("METHOD", "GENDER", "RACE", "ANY_SERVICE"), function(thename){
  x <- res_cat[[thename]]
  out <- x$parameters[, c("Class", "Category", "Prob")]
  out <- reshape(out, v.names = "Prob", timevar = "Class", idvar = "Category", direction = "wide")
  out$chisq <- NA
  out$p <- NA
  out$chisq[1] <- x$tests$Chi.square
  out$p[1] <- x$tests$p
  out <- cbind(Variable = NA, out)
  out$Variable[1] <- thename
  out$Category <- ct_dct[[thename]][out$Category]
  out
}))
cats$Variable <- rename_vars[cats$Variable]
cats[["p 1 vs 2"]] <- sapply(res_catdum, function(x){x$tests$p[1]})[match(cats$Category, dummie_dict)]
rownames(cats) <- NULL
knitr::kable(cats, digits = 2, caption = "Class differences in categorical auxiliary variables.")
```

>>>>

rename_vars <- c(
CDE7_NEW = "Told someone",
CDE8 = "Will likely attempt",
CDE9 = "Certain fatal",
CDE10 = "Certain lethal",
CDE11 = "Intended to die",
CDE12 = "Could be fixed",
CDE13 = "Attempts",
CDE16 = "Medical attention (DV)",
CDE17 = "NSSI",
AGE = "Age (cov)",
EDUCATION = "Education (cov)",
GENDER = "Gender (cov)",
ANY_SERVICE = "Service (cov)",
METHOD = "Method (DV)",
RACE = "Race (cov)",
CARE_CAT = "Medical attention cat (DV)"
)
df_plot$Variable <- ordered(df_plot$Variable, 
                            levels = unique(df_plot$Variable),
                            labels = rename_vars[unique(df_plot$Variable)])

@@

```{r mplusmodel}
df$care_cat <- factor(df$CDE16)
aux_cat <- c("Gender", "Race", "Any_Service", "Method", "care_cat")
aux_cont <- c("Age", "Education", "CDE16")
df_anal <- df[c(indicators, aux_cat, aux_cont)]
cat_dict <- lapply(df_anal[aux_cat], levels)
df_anal[aux_cat] <- lapply(df_anal[aux_cat], as.integer)
cat_vars <- c("CDE7_NEW", "CDE8",
              "CDE9", "CDE10", "CDE11", "CDE12", "CDE13", 
              "CDE17")

  # Create dummies for all categories to get by-category comparisons
  df_dum <- df_anal[c(indicators, aux_cat, aux_cont)]
  df_dum <- data.frame(lapply(df_dum, as.factor))
  dum_vars <- aux_cat[sapply(df_dum[aux_cat], function(x){length(unique(na.omit(x)))}) > 2]

  the_opt <- getOption("na.action")
  options(na.action='na.pass')
  dummies <- model.matrix(as.formula(paste0("~ -1 + ", paste0(dum_vars, collapse = " + "))), df_dum,
                          contrasts.arg = lapply(df_dum[dum_vars], contrasts, contrasts = FALSE))

  dummies <- data.frame(dummies, check.names = FALSE)
  options(na.action=the_opt)
  names(dummies)[nchar(names(dummies)) > 8] <- paste0(
    substr(names(dummies)[nchar(names(dummies)) > 8], 1, 7),
    sub('.*(?=.$)', '', names(dummies)[nchar(names(dummies)) > 8], perl=T)
  )

  # for(i in dum_vars){
  #   names(dummies) <- gsub(paste0("^", i, "(?!Other)"), "", names(dummies), perl = TRUE)
  # }
  #dummie_dict <- gsub("^.+?Other", "Other", names(dummies))
  #names(dummies) <- gsub("[ /\\(].*", "", names(dummies))
  df_dum <- cbind(df_dum, dummies)
if(run_everything){
  mod_cont <- mplusObject(
    TITLE = paste0("2 class with auxiliary continuous"),
    VARIABLE = paste0("CLASSES = c (2);\n",
                      paste0("CATEGORICAL = \n", paste0(cat_vars, collapse = "\n"), ";\n"),
                      paste0("USEVARIABLES = \n", paste0(indicators, collapse = "\n"), ";\n"),
                      paste0(c("AUXILIARY = \n",
                               paste0(aux_cont, " (BCH)\n"),
                               #paste0(aux_cat, " (DCATEGORICAL)\n"),
                               ";\n"), collapse = "")),
    ANALYSIS = "TYPE IS MIXTURE;\nLRTSTARTS = 0 0 40 10;\nOPTSEED = 637345;\n",
    OUTPUT = "TECH14;",
    SAVEDATA = paste0("FILE IS dat_2_aux_cont.dat;\nSAVE = CPROBABILITIES;"),
    usevariables = names(df_anal),
    rdata = df_anal
  )

  res_cont <- mplusModeler(mod_cont,
                           modelout = paste0("mod_2_aux_cont.inp"),
                           run = 1L)$results
  #df_anal$Any_Service <- droplevels()
  mod_cat <- mplusObject(
    TITLE = paste0("2 class with auxiliary categorical"),
    VARIABLE = paste0("CLASSES = c (2);\n",
                      paste0("CATEGORICAL = \n", paste0(cat_vars, collapse = "\n"), ";\n"),
                      paste0("USEVARIABLES = \n", paste0(indicators, collapse = "\n"), ";\n"),
                      paste0(c("AUXILIARY = \n",
                               #paste0(aux_cont, " (BCH)\n"),
                               paste0(aux_cat, " (DCATEGORICAL)\n"),
                               ";\n"), collapse = "")),
    ANALYSIS = "TYPE IS MIXTURE;\nLRTSTARTS = 0 0 40 10;\nOPTSEED = 637345;\n",
    OUTPUT = "TECH14;",
    SAVEDATA = paste0("FILE IS dat_2_aux_cat.dat;\nSAVE = CPROBABILITIES;"),
    usevariables = names(df_anal),
    rdata = df_anal
  )

@@

} else {
  res_cont <- readModels("mod_2_aux_cont.out")
  #res_cat <- extract_cat_tests(readLines("mod_2_aux_cat.out"))
  res_cat <- readModels("mod_2_aux_cat.out")$lcCondMeans
  res_catdum <- extract_cat_tests(readLines("mod_2_aux_catdum.out"))
  #names(res_catdum) <- dummie_dict
}
```

@@

```{r tabcont, include = TRUE}
tab <- res_cont$lcCondMeans$overall
tab$ci1 <- tidySEM::conf_int(tab$m.1, se = tab$se.1)
tab$ci2 <- tidySEM::conf_int(tab$m.2, se = tab$se.2)
tab$CramersV <- sqrt((as.numeric(tab$chisq)/res_cont$summaries$Observations)/1)
tab$var <- rename_vars[tab$var]
tab$df <- NULL
names(tab)[1] <- "Variable"
tab <- tab[c(grep("(DV)", tab$Variable), grep("(cov)", tab$Variable)), ]
rownames(tab) <- NULL
tab <- tab[, c("Variable", "m.1", "se.1", "ci1", "m.2", "se.2", "ci2","chisq", "p")]
knitr::kable(tab, digits = 2, caption = "Class differences in continuous auxiliary variables.")
```

```{r tabcat, include = TRUE}
ct_dct <- cat_dict
names(ct_dct) <- toupper(names(ct_dct))
#c("METHOD", "GENDER", "RACE", "ANY_SERVICE")
cats <- do.call(rbind, lapply(names(res_cat), function(thename){
  x <- res_cat[[thename]]
  out <- x$parameters[, c("class", "category", "prob", "se")]
  out$ci <- tidySEM::conf_int(as.numeric(out$prob), se = as.numeric(out$se))
  out <- reshape(out, v.names = c("prob", "se", "ci"), timevar = "class", idvar = "category", direction = "wide")
  out$chisq <- NA
  out$p <- NA
  out$cramersv <- NA
  out$chisq[1] <- x$tests$chi_square
  out$p[1] <- x$tests$p
  the_n <- x$tests$n
  if(is.null(the_n)) the_n <- res_cont$summaries$Observations
  out$cramersv[1] <- sqrt((as.numeric(x$tests$chi_square)/as.numeric(the_n))/1)
  out <- cbind(Variable = NA, out)
  out$Variable[1] <- thename
  out$category <- ct_dct[[thename]][out$category]
  out
}))
test_res <- sapply(res_catdum, function(x){x$tests$p[1]})
names_testres <- unique(gsub(".$", "", names(test_res)))
test_res <- lapply(names_testres, function(x){test_res[grep(x, names(test_res))]})
names(test_res) <- names_testres
#test_res <- test_res[na.omit(pmatch(na.omit(cats$Variable), names(test_res)))]

cats[["p 1 vs 2"]] <- NA
for(i in names(test_res)){
  #i = names(test_res)[1]
  cats$`p 1 vs 2`[grep(i, cats$Variable):(grep(i, cats$Variable) + (length(test_res[[i]])-1))] <- test_res[[i]]
}  
cats$Variable <- rename_vars[cats$Variable]
  #sapply(res_catdum, function(x){x$tests$p[1]})[match(cats$Category, dummie_dict)]
rownames(cats) <- NULL
knitr::kable(cats, digits = 2, caption = "Class differences in categorical auxiliary variables.")
```
